---
title: "Example usage"
data: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example usage}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

```{r setup}
library(GIST)
```

## Download the data

You can get some test data from [here]().

Create a folder called `test_data` and download the compressed data there. Unzip the compressed file. You should see the following folders:

```{r}
test_data = "~/Downloads/test_data"
list.files(test_data)
```

## Load and preprocess the ST data

Let's load the spatial transcriptomics (ST) data. 

```{r}
st_counts = load10xMatrix(file.path(test_data, "filtered_feature_bc_matrix"))
st_counts = preprocess_expr_mat(st_counts, impute = TRUE, k = 5, d = 10, normalize = 'sct')
```

## Load and precprocess the SC data

Let's load and normalize the SC data. Please note that for this dataset we need to pass `check.names` as `FALSE` to `read.table`, otherwise `R` will unhelpfully change the column names.

```{r}
sc_counts = read.table(file.path(test_data, "mouse-subsampled-scdata", "counts.tsv"), header = TRUE, row.names = 1, check.names = FALSE)
sc_counts = preprocess_expr_mat(sc_counts, impute = FALSE, normalize = 'sct')
```

## Make the signature matrix

We will now load the the labels for the cell types and construct the signature matrix.

```{r}
sc_labels = read.table(file.path(test_data, "mouse-subsampled-scdata", "cell_labels.tsv"), header = TRUE)
sig_mat = make_signature_matrix(sc_counts, sc_labels)
```

## Run GIST

```{r}
st_props = GIST(st_counts, sig_mat, num_cores = 1)
```

## Visualize results

```{r}
tp = load10xTP(file.path(test_data, "spatial", "tissue_positions_list.csv"))
plot_spots(st_props, tp, ncols = 3)
```
